---
title: "R Notebook"
output: html_notebook
---

Orpet, RJ, WR Cooper, EH Beers, and LB Nottingham. 2021. Test of plant defense elicitors for arthropod pest suppression and PR-1 gene induction in pear orchards. Entomologia Experimentalis et Applicata 169:1137-1146.

Data collection: Nottingham & Cooper
Code: Orpet

Figures for 2017 Experiment Data (Figs. 1 and 2)
```{r}
#LOAD DATA
library(readxl)
library(ggpubr)
Data <- read_excel("1728 analysis - leaf brush.xlsx")


#REORDER TREATMENT LABELS
Data$trt <- factor(Data$trt, levels=c("Control", "Actigard", "Employ"), labels=c("Control", "ASM", "Harpin"))
Data$site <- factor(Data$site,
                    levels=c("A","B","C","D","E","F"),
                    labels=c("Orchard 1", "Orchard 2", "Orchard 3","Orchard 4", "Orchard 5", "Orchard 6"))


library(ggplot2)

#Pear psylla eggs panel
eggs <- ggplot(data = Data, aes(x = day, y = eggs/50, shape = trt, linetype = trt, color = trt, fill = trt)) +
  stat_summary(fun.data = mean_se, geom="errorbar", width=0,  size=0.4, position=position_dodge(3), linetype = 'solid') +
  stat_summary(fun = mean, geom = "line", position=position_dodge(3)) +
  stat_summary(fun = mean, geom = "point", position=position_dodge(3)) +
  scale_color_manual(values=c('black', 'black', 'grey')) +
  scale_shape_manual(values=c(21, 23, 23)) +
  scale_linetype_manual(values = c('dotted', 'solid', 'solid')) +
  scale_fill_manual(values=c('white', 'black', 'grey')) +
  scale_x_continuous(limits = c(207,245), breaks = c(212, 227, 243), labels = c('31 Jul', '15 Aug', '31 Aug')) +
  #scale_y_continuous(expand = expansion(mult = c(0.05, 0))) +
  facet_wrap(~site, scales = "free_y") +
  ylab("No. eggs/leaf\n") +
  xlab("\nDate") +
  #ggtitle("A") +
  guides(fill=guide_legend(ncol=1)) +
  theme_classic() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 0.75) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 0.75) +
  theme(
        strip.text = element_text(vjust = -1, size = 8),
        plot.title = element_text(size = 11, face = "bold"),
        strip.background = element_blank(),
        #panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        axis.line = element_line(colour = 'black', size = 0),
        axis.title.x = element_blank(),
        axis.ticks = element_line(colour = "black"),
        axis.text.x = element_text(color = 'black'),
        axis.text.y = element_text(color = 'black'),
        plot.margin = (unit(c(1,1,0,1),"mm")),
                #transparent background
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_rect(fill = "transparent",colour = NA),
                plot.background = element_rect(fill = "transparent",colour = NA),
                legend.position=c(1,1), legend.justification=c(1,0),legend.box="vertical", legend.margin=margin(), legend.title = element_blank(), legend.key.size = unit(0.75, 'lines')
        )
  
#Pear psylla nymphs panel
nymphs <- ggplot(data = Data, aes(x = day, y = (old+yng)/50, shape = trt, linetype = trt, fill = trt, color = trt)) +
  stat_summary(fun.data = mean_se, geom="errorbar", width=0,  size=0.4, position=position_dodge(3), linetype = 'solid') +
  stat_summary(fun = mean, geom = "line", position=position_dodge(3)) +
  stat_summary(fun = mean, geom = "point", position=position_dodge(3)) +
  scale_color_manual(values=c('black', 'black', 'grey')) +
  scale_shape_manual(values=c(21, 23, 23)) +
  scale_linetype_manual(values = c('dotted', 'solid', 'solid')) +
  scale_fill_manual(values=c('white', 'black', 'grey')) +
  scale_x_continuous(limits = c(207,245), breaks = c(212, 227, 243), labels = c('31 Jul', '15 Aug', '31 Aug')) +
  #scale_y_continuous(expand = expansion(mult = c(0.05, 0))) +
  facet_wrap(~site, scales = "free_y") +
  ylab("No. nymphs/leaf\n") +
  xlab("\nDate") +
  #ggtitle("B") +
  guides(fill=guide_legend(ncol=1)) +
  theme_classic() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 0.75) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 0.75) +
  theme(
        strip.text = element_text(vjust = -1, size = 8),
        plot.title = element_text(size = 11, face = "bold"),
        strip.background = element_blank(),
        #panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        axis.line = element_line(colour = 'black', size = 0),
        #axis.title.x = element_blank(),
        axis.ticks = element_line(colour = "black"),
        axis.text.x = element_text(color = 'black'),
        axis.text.y = element_text(color = 'black'),
        plot.margin = (unit(c(0,0,0,0),"mm")),
                #transparent background
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_rect(fill = "transparent",colour = NA),
                plot.background = element_rect(fill = "transparent",colour = NA),
                legend.position=c(1,1), legend.justification=c(1,0),legend.box="vertical", legend.margin=margin(), legend.title = element_blank(), legend.key.size = unit(0.75, 'lines')
        )
  


output <- ggarrange(
    eggs,
    nymphs,
    ncol = 1,
    align="hv", common.legend = TRUE)
  #ggsave(plot = output, file = paste("file",Site,".png",sep=""), width = 8.5, height = 10, path = "outputgraphs")

ggsave(plot = output, file = "Figure_1.tiff", width = 16, height = 14, units = "cm", dpi = 400)


#Spider mites panel
Data <- read_excel("1728 analysis - leaf brush.xlsx")
Data$trt <- factor(Data$trt, levels=c("Control", "Actigard", "Employ"), labels=c("Control", "ASM", "Harpin"))
  #Exclude sites A and B due to very low counts
  Data <- subset(Data, site != "A")
  Data <- subset(Data, site != "B")
  Data$site <- factor(Data$site,
                    levels=c("C","D","E","F"),
                    labels=c("Orchard 3","Orchard 4", "Orchard 5", "Orchard 6"))

output <- ggplot(data = Data, aes(x = day, y = tsm/50, shape = trt, linetype = trt, color = trt, fill = trt)) +
  stat_summary(fun.data = mean_se, geom="errorbar", width=0,  size=0.4, position=position_dodge(3), linetype = 'solid') +
  stat_summary(fun = mean, geom = "line", position=position_dodge(3)) +
  stat_summary(fun = mean, geom = "point", position=position_dodge(3)) +
  scale_color_manual(values=c('black', 'black', 'grey')) +
  scale_shape_manual(values=c(21, 23, 23)) +
  scale_linetype_manual(values = c('dotted', 'solid', 'solid')) +
  scale_fill_manual(values=c('white', 'black', 'grey')) +
  scale_x_continuous(limits = c(207,245), breaks = c(212, 227, 243), labels = c('31 Jul', '15 Aug', '31 Aug')) +
  #scale_y_continuous(expand = expansion(mult = c(0.05, 0))) +
  facet_wrap(~site, scales = "free_y", ncol = 1) +
  ylab("No. spider mites/leaf\n") +
  xlab("\nDate") +
  guides(fill=guide_legend(nrow = 1)) +
  theme_classic() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 0.75) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 0.75) +
  theme(
        strip.text = element_text(vjust = -1, size = 8),
        plot.title = element_text(size = 11, face = "bold"),
        strip.background = element_blank(),
        #panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        #axis.line = element_line(colour = 'black', size = 0),
        #axis.title.x = element_blank(),
        axis.ticks = element_line(colour = "black"),
        axis.text.x = element_text(color = 'black'),
        axis.text.y = element_text(color = 'black'),
                #transparent background
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_rect(fill = "transparent",colour = NA),
                plot.background = element_rect(fill = "transparent",colour = NA),
                legend.position=c(1,1), legend.justification=c(1,0),legend.box="vertical", legend.margin=margin(), legend.title = element_blank(), legend.key.size = unit(0.75, 'lines')
        )
output <- ggarrange(
    output,
    ncol = 1,
    align="hv", common.legend = TRUE)
  
ggsave(plot = output, file = "Figure_2.tiff", width = 6, height = 11, units = "cm", dpi = 400)


```



PR-1 relative expression (Figure 4)
```{r}

library(readxl)
library(ggplot2)
Data <- read_excel("Fig 4.xlsx")
Data$Weeks <- as.character(Data$Weeks)
Data$Experimentlabel <- factor(Data$Experiment, labels=c(
               "A"=expression(paste("27 April field spray")),
               "B"=expression(paste("23 May field spray")),
               "C"=expression(paste("Potted tree spray "))))

ggplot(data = Data, aes(x = Weeks, y = minusddCT)) +
  #geom_errorbar(aes(ymin = Foldchange-SEM, ymax = Foldchange+SEM, group = Elicitor), width = 0, position=position_dodge(width=0.9)) +
  #geom_point(stat = "identity", shape = 21, color = 'black', aes(fill = Elicitor), position=position_dodge(width=0.5)) +
  geom_hline(yintercept = 0, linetype = 'dashed', size = 0.25) +
  geom_point(shape = 21, size = 2, aes(group = Elicitor), position=position_dodge(width=0.6)) +
  geom_errorbar(stat = "summary", fun.data = mean_se, aes(group = Elicitor), position=position_dodge(width=0.6), width=0) +
  geom_point(shape = 23, size = 2, stat = "summary", fun = "mean", color = 'black', aes(fill = Elicitor), position=position_dodge(width=0.6)) +
  facet_wrap(~Experimentlabel, ncol = 1, scales = 'free',labeller = label_parsed) +
  scale_fill_manual(values=c('black', 'grey')) +
  geom_blank(data = subset(Data, Experiment  == "A"), aes(y = 3)) +
  geom_blank(data = subset(Data, Experiment  == "B"), aes(y = 4, x = '1')) +
  geom_blank(data = subset(Data, Experiment  == "C"), aes(x = '7')) +
  scale_y_continuous(expression(paste("Relative ",italic("PR-1 "), "expression (-",paste(Delta),paste(Delta),"C"[T],")"))) +
  scale_x_discrete("\nWeeks post-treatment",breaks = c("0 (pre-spray)","1","3","7")) +
  theme_classic() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 0.75) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 0.75) +
  guides(fill=guide_legend(ncol=1)) +
  theme(
        #strip.text = element_blank(),
        #strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        #panel.border = element_rect(colour = "black", fill = NA),
        axis.line = element_line(colour = 'black', size = 0.5),
        axis.ticks = element_line(colour = "black"),
        axis.text.x = element_text(color = 'black'),
        axis.text.y = element_text(color = 'black'),
        strip.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.position=c(1,1), legend.justification=c(1,0),legend.box="vertical", legend.margin=margin(), legend.title = element_blank(), legend.key.size = unit(0.75, 'lines')
        #panel.spacing = unit(1.5, "lines")
        )

ggsave("Figure_4.tiff", width = 8, height = 12, units = "cm", dpi = 400, bg = "transparent")

  

```


Statistical Analysis for PR-1 data (Fig 4)
```{r}
#LOAD PACKAGES
library(readxl)
library(car) #for Anova
library(glmmTMB) #for generalized linear models
library(DHARMa) #for model diagnostics
library(dplyr) #for filter()
library(emmeans) #for tukey tests


#LOAD DATA
Data <- read_excel("Fig 4.xlsx")

#convert week to categorical variable
Data$Weeks <- as.factor(Data$Weeks)

#BUILD MODELS
for (test in unique(Data$Experiment)) {
  Data_test <- subset(Data, Experiment == test)
  for(pde in unique(Data_test$Elicitor)){
  Data_test_by_elicitor <- subset(Data_test, Elicitor == pde)
  model <- glmmTMB(formula = minusddCT ~ Weeks + (1|Block), data = Data_test_by_elicitor)
  plot(simulateResiduals(model))
  print(test)
  print(pde)
  print(Anova(model))
  }
}

print(emmeans(model, trt.vs.ctrl ~ source ~ Weeks,adjust = "mvt"))


```


Statistical Analysis for 2017 Experiment (Table 3)
```{r}
#LOAD PACKAGES
library(readxl)
library(car) #for Anova
library(glmmTMB) #for generalized linear models
library(DHARMa) #for model diagnostics
library(dplyr) #for filter()
library(emmeans) #for tukey tests


#LOAD DATA
Data <- read_excel("1728 analysis - leaf brush.xlsx")

#convert week to categorical variable
Data$week <- as.factor(Data$week)

#BUILD MODELS
eggsmodel   <- glmmTMB(formula = eggs    ~ trt + week + trt:week + site + trt:site + (1|site:block) + (1|subject), family = "nbinom1", data = Data)
nymphsmodel <- glmmTMB(formula = old+yng ~ trt + week + trt:week + site + trt:site + (1|site:block) + (1|subject), family = "nbinom1", data = Data)
tsmmodel    <- glmmTMB(formula = tsm     ~ trt + week + trt:week + site + trt:site + (1|site:block) + (1|subject), family = "nbinom2",
                       data = subset(Data, site != 'A' & site != 'B')) #exclude blocks A and B due to low tsm counts

#MODEL DIAGNOSTICS
plot(simulateResiduals(eggsmodel))
plot(simulateResiduals(nymphsmodel))
plot(simulateResiduals(tsmmodel))

#ANOVAs
Anova(eggsmodel)
Anova(nymphsmodel)
Anova(tsmmodel)

```



Figures for 2018 Experiment Data (Fig. 3)
```{r}

####Note: There are samples for June 12 and 13. June 12 (and all previous samples) used dishwashing liquid on the plates. June 13 and following did not use dishwashing liquid because it made it difficult to tell whether nymphs were alive. Because of the closeness in time of the June 12 and 13 samples, only the June 13 data were used to represent that week.


####LOAD DATA####
library(readxl)
library(ggplot2)
Data <- read_excel("1828 - Analysis - Leaf brush.xlsx")
Data <- subset(Data, day != 163) # exclude the June 12 sample explained above

#reorder trt labels
Data$trt <- factor(Data$trt, levels=c("Control", "Actigard", "Employ"), labels=c("Control", "ASM", "Harpin"))
Data$Timing <- factor(Data$Timing, levels=c("1", "2"), labels=c("i) 27 April", "ii) 23 May"))



eggs <- ggplot(data = Data, aes(x = day, y = ((eggs/lvs)), shape = trt, linetype = trt, color = trt, fill = trt)) +
  stat_summary(fun.data = mean_se, geom="errorbar", width=0,  size=0.4, position=position_dodge(5), linetype = 'solid') +
  stat_summary(fun.y = mean, geom = "line", position=position_dodge(5)) +
  stat_summary(fun.y = mean, geom = "point", position=position_dodge(5)) +
  scale_color_manual(values=c('black', 'black', 'grey')) +
  scale_shape_manual(values=c(21, 23, 23)) +
  scale_linetype_manual(values = c('dotted', 'solid', 'solid')) +
  scale_fill_manual(values=c('white', 'black', 'grey')) +
  scale_x_continuous(limits = c(110,200), breaks = c(117, 155, 193), labels = c('17\nApr', '4\nJun', ' 12\nJul')) +
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3)) +
  facet_wrap(~Timing, ncol = 1) +
  ylab("No. eggs/leaf") +
  xlab("\nDate") +
  #ggtitle("A") +
  theme_classic() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 0.75) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 0.75) +
  theme(
        strip.text = element_blank(),
        plot.title = element_text(size = 11, face = "bold"),
        strip.background = element_blank(),
        #panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        axis.line = element_line(colour = 'black', size = 0),
        axis.title.x = element_blank(),
        axis.ticks = element_line(colour = "black"),
        axis.text.x = element_text(color = 'black'),
        axis.text.y = element_text(color = 'black'),
        #plot.margin = (unit(c(1,2,1,1),"mm")),
                #transparent background
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_rect(fill = "transparent",colour = NA),
                plot.background = element_rect(fill = "transparent",colour = NA),
                legend.position=c(1,1), legend.justification=c(1,0),legend.box="vertical", legend.margin=margin(), legend.title = element_blank(), legend.key.size = unit(0.75, 'lines')
        )
  

nymphs <- ggplot(data = Data, aes(x = day, y = (((yng+old)/lvs)), shape = trt, linetype = trt, color = trt, fill = trt)) +
  stat_summary(fun.data = mean_se, geom="errorbar", width=0,  size=0.4, position=position_dodge(5), linetype = 'solid') +
  stat_summary(fun.y = mean, geom = "line", position=position_dodge(5)) +
  stat_summary(fun.y = mean, geom = "point", position=position_dodge(5)) +
  scale_color_manual(values=c('black', 'black', 'grey')) +
  scale_shape_manual(values=c(21, 23, 23)) +
  scale_linetype_manual(values = c('dotted', 'solid', 'solid')) +
  scale_fill_manual(values=c('white', 'black', 'grey')) +
  scale_x_continuous(limits = c(110,200), breaks = c(117, 155, 193), labels = c('17\nApr', '4\nJun', ' 12\nJul')) +
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3)) +
  facet_wrap(~Timing, ncol = 1) +
  ylab("No. nymphs/leaf") +
  xlab("\nDate") +
  #ggtitle("B") +
  theme_classic() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 0.75) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 0.75) +
  theme(
        strip.text = element_blank(),
        plot.title = element_text(size = 11, face = "bold"),
        strip.background = element_blank(),
        #panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        axis.line = element_line(colour = 'black', size = 0),
        #axis.title.x = element_blank(),
        axis.ticks = element_line(colour = "black"),
        axis.text.x = element_text(color = 'black'),
        axis.text.y = element_text(color = 'black'),
        #plot.margin = (unit(c(1,2,1,1),"mm")),
                #transparent background
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_rect(fill = "transparent",colour = NA),
                plot.background = element_rect(fill = "transparent",colour = NA),
                legend.position=c(1,1), legend.justification=c(1,0),legend.box="vertical", legend.margin=margin(), legend.title = element_blank(), legend.key.size = unit(0.75, 'lines')
        )
  

#Adult taps
Data <- read_excel("1828 - Analysis - Taps.xlsx")

#reorder trt labels
Data$trt <- factor(Data$trt, levels=c("Control", "Actigard", "Employ"), labels=c("Control", "ASM", "Harpin"))

#First, need to average taps within reps together

summary1 <- do.call(data.frame, 
                        aggregate(adults~Timing+day+trt+block, 
                                  data = Data, 
                                  function(x) 
                                    c(mean = mean(x),
                                      sum = sum(x),
                                      sd = sd(x), 
                                      count = length(x))))

summary1$Timing <- factor(summary1$Timing, levels=c("1", "2"), labels=c("i) 27 April", "ii) 23 May"))

adults <- ggplot(data = summary1, aes(x = day, y = adults.mean, shape = trt, linetype = trt, color = trt, fill = trt)) +
  stat_summary(fun.data = mean_se, geom="errorbar", width=0,  size=0.4, position=position_dodge(5), linetype = 'solid') +
  stat_summary(fun.y = mean, geom = "line", position=position_dodge(5)) +
  stat_summary(fun.y = mean, geom = "point", position=position_dodge(5)) +
  scale_color_manual(values=c('black', 'black', 'grey')) +
  scale_shape_manual(values=c(21, 23, 23)) +
  scale_linetype_manual(values = c('dotted', 'solid', 'solid')) +
  scale_fill_manual(values=c('white', 'black', 'grey')) +
  scale_x_continuous(limits = c(110,200), breaks = c(117, 155, 193), labels = c('17\nApr', '4\nJun', ' 12\nJul')) +
  #scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3)) +
  facet_wrap(~Timing, ncol = 1) +
  ylab("No. adults/tap") +
  xlab("\nDate") +
  #ggtitle("C") +
  theme_classic() +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 0.75) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 0.75) +
  theme(
        strip.text = element_blank(),
        plot.title = element_text(size = 11, face = "bold"),
        strip.background = element_blank(),
        #panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        axis.line = element_line(colour = 'black', size = 0),
        axis.title.x = element_blank(),
        axis.ticks = element_line(colour = "black"),
        axis.text.x = element_text(color = 'black'),
        axis.text.y = element_text(color = 'black'),
        #plot.margin = (unit(c(1,1,1,2),"mm")),
                #transparent background
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_rect(fill = "transparent",colour = NA),
                plot.background = element_rect(fill = "transparent",colour = NA),
                legend.position=c(1,1), legend.justification=c(1,0),legend.box="vertical", legend.margin=margin(), legend.title = element_blank(), legend.key.size = unit(0.75, 'lines')
        )
  


output <- ggarrange(
    eggs,
    nymphs,
    adults,
    ncol = 3,
    align="hv", common.legend = TRUE)
  
ggsave(plot = output, file = "Figure_3.tiff", width = 16, height = 6, units = "cm", dpi = 400)




```

Statistical models for 2018 experiment
```{r}

#Note: There are samples for June 12 and 13. June 12 (and all previous samples) used dishwashing liquid on the plates. June 13 and following did no use dishwashing liquid because it made it difficult to tell whether nymphs were alive. Because of the closeness in time of the June 12 and 13 samples, only the June 13 data were used.

#LOAD PACKAGES
library(readxl)
library(ggplot2)
library(glmmTMB)
library(DHARMa)
library(car)
library(emmeans)

#LOAD DATA
#Nymph data
Data <- read_excel("1828 - Analysis - Leaf brush.xlsx")
Data <- subset(Data, day != 163) # exclude the June 12 sample explained above
Data$week <- as.factor(Data$week)
Data$Timing <- as.factor(Data$Timing)
Dataspray1 <- subset(Data, Timing == 1)
Dataspray2 <- subset(Data, Timing == 2)
#Adult tap data
Data_taps <- read_excel("1828 - Analysis - Taps.xlsx")
Data_taps$week <- as.factor(Data_taps$week)
Data_taps$Timing <- as.factor(Data_taps$Timing)
Data_taps <- do.call(data.frame, 
                          aggregate(adults~Timing+day+week+trt+block+subject, 
                                    data = Data_taps, 
                                    function(x) 
                                    c(sum = sum(x), 
                                      count = length(x))))
Dataspray1_taps <- subset(Data_taps, Timing == 1)
Dataspray2_taps <- subset(Data_taps, Timing == 2)

#BUILD MODELS
eggsmodel_spray1   <- glmmTMB(formula = log((eggs/lvs)+0.01)  ~ trt + week + trt:week + (1|block) + (1|subject), family = "gaussian", data = Dataspray1)
eggsmodel_spray2   <- glmmTMB(formula = log((eggs/lvs)+0.01)  ~ trt + week + trt:week + (1|block) + (1|subject), family = "gaussian", data = Dataspray2)
nymphsmodel_spray1 <- glmmTMB(formula = log((nymphs/lvs)+0.01)  ~ trt + week + trt:week + (1|block) + (1|subject), family = "gaussian", data = Dataspray1)
nymphsmodel_spray2 <- glmmTMB(formula = log((nymphs/lvs)+0.01)  ~ trt + week + trt:week + (1|block) + (1|subject), family = "gaussian", data = Dataspray2)
adultmodel_spray1  <- glmmTMB(formula = log(adults.sum+1)  ~ trt + week + trt:week + (1|block) + (1|subject), family = "gaussian", data = Dataspray1_taps)
adultmodel_spray2  <- glmmTMB(formula = log(adults.sum+1)  ~ trt + week + trt:week + (1|block) + (1|subject), family = "gaussian", data = Dataspray2_taps)

#MODEL DIAGNOSTICS
plot(simulateResiduals(eggsmodel_spray1))
plot(simulateResiduals(eggsmodel_spray2))
plot(simulateResiduals(nymphsmodel_spray1))
plot(simulateResiduals(nymphsmodel_spray2))
plot(simulateResiduals(adultmodel_spray1))
plot(simulateResiduals(adultmodel_spray2))

#STATISTICAL TESTS
Anova(eggsmodel_spray1)
Anova(eggsmodel_spray2)
Anova(nymphsmodel_spray1)
Anova(nymphsmodel_spray2)
Anova(adultmodel_spray1)
Anova(adultmodel_spray2)

#POST-HOC TESTS FOR NYMPHS
#____
#conduct a Tukey test to compare treatment means within each sample date
#purpose of this is to diagnose the significant Week X Trt interactions for nymphs for both spray timing experiments
#note: poorly fitting models occur only for days with very low counts across treatments and are ignored as those data have too low counts for any practical interpretation anyway.
#____
#tests for spray experiment 1
dates <- unique(Dataspray1$day)
for (value in dates){
  analysisdata <- subset(Dataspray1, day == value)
  nymphsmodel1   <- glmmTMB(formula = log(nymphs/lvs+0.01)  ~ trt + (1|block), family = "gaussian", data = analysisdata)
print(paste("Day",value,"readout begins:"))
plot(simulateResiduals(nymphsmodel1))
print(emmeans(nymphsmodel1, list(pairwise ~ trt), adjust = "tukey"))
}
#tests for spray experiment 2
dates <- unique(Dataspray2$day)
for (value in dates){
  analysisdata <- subset(Dataspray2, day == value)
  nymphsmodel1   <- glmmTMB(formula = log(nymphs/lvs+0.01)  ~ trt + (1|block), family = "gaussian", data = analysisdata)
print(paste("Day",value,"readout begins:"))
plot(simulateResiduals(nymphsmodel1))
print(emmeans(nymphsmodel1, list(pairwise ~ trt), adjust = "tukey"))
}

```

